---
title: "Latex"
subtitle: "Editing for Mathematics"
author: "Massimiliano Porto"
date: '`r format(Sys.time(), "%d %m, %Y")`'
output: 
  pdf_document:
    toc: yes
    number_sections: true
---

# Greek letters

Here a list of some of the most used Greek letters in Economics:

```{r, include = FALSE}
library(tidyverse)

# "sin" -> "\\sin"
latexify = function(base, asis = FALSE) {
  if (asis) return(base)
  paste0("\\", base)
}

# "sin" -> "`\\sin`" 
put_codechunk = function(base, ...) paste0("`", latexify(base, ...), "`")

# "sin" -> "$\\sin$"
put_result = function(base, ...) paste0("$", latexify(base, ...), "$")

# Create a cheatsheet table
math_table = function(cmd, description = NULL) {
  # cmd = c("sin", "cos", "log") 
  # description = c("sine", "cosine", "logarithm")
  #
  # => A tibble with additional columns like 
  #      latex = c("\\sin", "\\cos", "\\log") 
  #    will be created 
  # If description is NULL, cmd column is copied. 
  
  if (is.null(description)) description = cmd 
  tbl = tibble(description, cmd) %>% 
    mutate(latex = put_codechunk(cmd), 
           result = put_result(cmd)) %>% 
    select(-cmd)
}

two_column_layout = function(s) {
  cat("<div class=\"twocol\">")
  cat("<div class=\"left\">")
  cat("**LaTeX CODE**\n")
  cat("```", "$$", s, "$$", "```", sep = "\n")
  cat("</div><!-- left -->")
  cat("<div class=\"right\">")
  cat("**RESULT**\n")
  cat("$$", s, "$$", sep = "\n")
  cat("</div><!-- right -->")
  cat("</div><!-- twocol -->")
}
```


```{r, echo = FALSE}
greek_letters = c('alpha', 'beta', 'gamma', 'theta', 'sigma', 
                  'rho', 'lambda', 'phi', 'psi', 'epsilon')
knitr::kable(math_table(greek_letters))
```

For the corresponding capital letter, we just write in the same way but with initial capital letter.

Example

```{r, echo = FALSE}
capital_letters = c("Gamma", "Lambda", "Sigma", "Psi", 
                    "Delta", "Xi", "Upsilon", "Omega", 
                    "Theta", "Pi", "Phi")
knitr::kable(math_table(capital_letters))
```


#Parenthesis
We can write parenthesis and brackets just giving the input on the keyboard $()$ or $[]$. This does not work for braces.

Another way is to write parenthesis with the following notation:

```{r, echo = FALSE, results = "asis"}
s_paren = paste(
  "\\left(",
  "\\right) \\\\",
  "\\left[",
  "\\right]", 
  sep = "\n") %>% two_column_layout
```


that is writing `\left` _before opening the parenthesis_ and `\right` _before closing the parenthesis_. 

In case of braces, we should use the following notation:

```{r, echo = FALSE, results = "asis"}
s_curly = paste(
  "\\left\\{", 
  "\\right\\}",
  sep = "\n") %>% two_column_layout
```

Note the difference with brackets that are written `\left\` and `\right\`.

Example:

```{r, echo = FALSE, results = "asis"}
s_complex_braces = paste(
"10 \\cdot ",
"  \\left\\{",
"    5 + \\left[",
"      3 + 2 \\cdot ",
"         \\left(",
"            9-1\\right)",
"        \\right]  ",
"  \\right\\}",
sep = "\n") %>% two_column_layout
```

The opportunity to write the full notation for parenthesis and brackets will be clear when dealing whit fractions. 

#Fractions
To write fractions the inputs to give are `\frac` and then _numerator and denominator written between two braces_. With an example will be clear.

Example:

```{r, echo = FALSE, results = "asis"}
s_frac = paste(
"\\frac{5}{2}",
sep = "\n")%>% two_column_layout

```


Example:

```{r, echo = FALSE, results = "asis"}
s_frac01 = paste(
"\\frac{\\alpha}{1-\\alpha} = \\phi",
sep = "\n")%>% two_column_layout

```



#Paranthesis & Fraction
If we write a fraction in parenthesis with the shortcut

```{r, echo=FALSE, results="asis"}
s_pfrac = paste(
 "( \\frac{5}{2})", 
sep = "\n")%>% two_column_layout

```

the parenthesis do not fit the fraction.

If we write with full notation

```{r, echo=FALSE, results="asis"}
s_pfrac01 = paste(
  "\\left(",
  "\\frac{5}{2}",
  "\\right)",
sep = "\n")%>% two_column_layout

```

we will have an appropriate result.

Example:

```{r, echo=FALSE, results="asis"}
s_pfrac02 = paste(
  "\\frac{Y}{L} =",
  "\\left(",
  "\\frac{K}{L}, 1",
  "\\right)",
sep = "\n")%>% two_column_layout

```



#Exponential

To write the exponential we use the symbol `^` 

```{r, echo=FALSE, results="asis"}
s_expo = paste(
"2^5",
sep = "\n")%>% two_column_layout

```


In case of operation in the exponential, we need to _use braces after_ `^` 


Example:

```{r, echo=FALSE, results="asis"}
s_expo01 = paste(
 "F(K, L) = ",
 "K^\\alpha L^\\beta",
sep = "\n")%>% two_column_layout

```


Example:

```{r, echo=FALSE, results="asis"}
s_expo02 = paste(
  "F(K, L) =",
  "\\left[",
  "\\gamma K^\\frac{\\sigma - 1}{\\sigma}",
  " + (1 - \\gamma)L^\\frac{\\sigma - 1}{\\sigma}",
  "\\right]^\\frac{\\sigma}{\\sigma - 1}",
sep="\n")%>% two_column_layout

```



#Subscript

To write subscript use the symbol `_`

```{r, echo=FALSE, results="asis"}
s_sub = paste(
  "t_0",
sep= "\n")%>% two_column_layout

```



We use _braces in case of operation in subscript_

Example

```{r, echo=FALSE, results="asis"}
s_sub01 = paste(
  "y_ty_{t-j}",
sep = "\n")%>% two_column_layout
```


#Derivative

Example:

```{r, echo=FALSE, results="asis"}
s_der = paste(
  "\\frac{dk}{k}",
sep="\n")%>% two_column_layout
```

To write partial differentiation, write `\partial` in the fraction notation

Example:

```{r, echo=FALSE, results="asis"}
s_der01 = paste(
  "\\frac{",
  "\\partial F}{",
  "\\partial K}",
  "r + \\delta",
sep="\n")%>% two_column_layout
```


#Summation

The code to write summation is the following: symbol of `\sum_` followed by _braces with lower bound of summation_ and `^` _before the brace with the upper bound of summation_. 

Example:

```{r, echo=FALSE, results="asis"}
s_sum = paste(
  "\\sum_{t_0}^{\\infty}",
sep = "\n")%>% two_column_layout
```

An example summing up what we saw until now:

```{r, echo=FALSE, results="asis"}
s_sum01 = paste(
  "\\mathcal L =",
  "\\sum_{t=0}^{\\infty}",
  "\\left(",
  "\\frac {1}{1 + \\rho}",
  "\\right)^{t} u(c_t)",
  "\\frac{L_t}{H}",
  " + \\lambda",
  "\\left[",
  "S_0 + \\sum_{t=0}^{\\infty}",
  "R_t^{-1}w_t",
  "\\left(",
  "\\frac{L_t}{H}",
  "\\right)",
  "- \\sum_{t=0}^{\\infty} R_t^{-1}c_t",
  "\\left(",
  "\\frac{L_t}{H}",
  "\\right)",
  "\\right]",
sep="\n")%>% two_column_layout
```



#Limit

The code for limit is written as follows:

```{r, echo = FALSE, results="asis"}
s_lim = paste(
  "\\lim_{t\\to\\infty}",
sep = "\n")%>% two_column_layout
```


Example: 

```{r, echo=FALSE, results="asis"}
s_lim01 = paste(
  "\\lim_",
  "{t\\to\\infty}",
  "R_t^{-1}S_t = 0",
sep = "\n")%>% two_column_layout
```

Example

```{r, echo = FALSE, results="asis"}
s_lim02 = paste(
  "e^{",
  "lim_{\\gamma\\to0}",
  "\\frac{1}{\\gamma}ln",
  "\\left[",
  "\\alpha K^\\gamma + (1 - \\alpha)L^\\gamma",
  "\\right]}",
sep = "\n")%>% two_column_layout
```


#Integrals

```{r, echo = FALSE, results="asis"}
s_int = paste(
  "\\int_0^\\infty",
sep = "\n")%>% two_column_layout
```


Example

<div class="twocol">
<div class="left">
```
$$
\int_0^\infty
  \left[
    r(t)e^{-R(t)}S(t) 
    + e^{-R(t)} w(t) 
    \frac{L(t)}{H}
    -e^{-R(t)}c(t) 
    \frac{L(t)}{H} 
  \right]
  dt
$$
```
</div><!-- .left -->
<div class="right">
$$
\int_0^\infty\left[r(t)e^{-R(t)}S(t) + e^{-R(t)} w(t) \frac{L(t)}{H}-e^{-R(t)}c(t) \frac{L(t)}{H} \right]dt
$$
</div><!-- .left -->
</div>

#Optimization problem

Following how we set an optimization problem:
```{r, echo=FALSE, results="asis"}
s_int = paste(
  "\\max_{x_{1}, x_{2}}", 
  "(x_{1}^{\\alpha} + x_{2}^{\\alpha})",
  "^{\\frac{1}{ \\alpha}}",
sep = "\n")%>% two_column_layout
```
```{r, echo=FALSE, results="asis"}
s_int = paste(
  "\\text {subject to}", 
  "\\quad p_{1}x_{1}+p_{2}x_{2} \\leq I",
sep = "\n")%>% two_column_layout
```

```{r, echo=FALSE, results="asis"}
s_int = paste(
  "\\mathcal L", 
  "(x_1, x_2, \\lambda) \\equiv",
sep = "\n")%>% two_column_layout
```
  
#Matrix

The following example shows how to write a matrix.

Note that `\bmatrix` stands for _bracket matrix_ while `\pmatrix` stands for _parenthesized matrix_. Other kinds of matrix are available.

Moreover, `\vdots` stands for _vertical dots_ while `\ddots` for _diagonal dots_. `\cdots` stands for _central dots_, used also as multiplication sign.

**LaTeX Code**
<div class="twocol">
<div class="left">
```
$$
A =
\begin{bmatrix}
  a_{11} & a_{12} & a_{13} & \dots & a_{1n} \\
  a_{21} & a_{22} & a_{23} & \dots & a_{2n} \\
  \vdots & \vdots & \vdots & \vdots & \vdots\\
  a_{m1} & a_{m2} & a_{m3} & \dots & a_{mn} \\
\end{bmatrix}
\ne
B =
\begin{pmatrix}
 b_{11} & b_{12} & b_{13} & \cdots & b_{1n} \\
 b_{21} & \ddots & b_{23} & \cdots & b_{2n} \\
 ...&...&... & \ddots &... \\
 b_{m1} & b_{m2} & b_{m3} & \dots & b_{mn} \\
\end{pmatrix}
$$
```
**RESULT**
</div><!-- .left -->
<div class="right">
$$
A =
\begin{bmatrix}
  a_{11} & a_{12} & a_{13} & \dots & a_{1n} \\
  a_{21} & a_{22} & a_{23} & \dots & a_{2n} \\
  \vdots & \vdots & \vdots & \vdots & \vdots\\
  a_{m1} & a_{m2} & a_{m3} & \dots & a_{mn} \\
\end{bmatrix}
\ne
B =
\begin{pmatrix}
 b_{11} & b_{12} & b_{13} & \cdots & b_{1n} \\
 b_{21} & \ddots & b_{23} & \cdots & b_{2n} \\
 ...&...&... & \ddots &... \\
 b_{m1} & b_{m2} & b_{m3} & \dots & b_{mn} \\
\end{pmatrix}
$$
</div><!-- .left -->
</div>


Another example with matrix

**LaTeX Code**
<div class="twocol">
<div class="left">
```
$$
PAQ=\left[
  \begin{array}{c|c}
    I_{d\times d} & O_{d\times(n-d)}\\
    \hline O_{(m-d)\times d} & O_{(m-d)\times(n-d)}
  \end{array}
\right]
$$
```
**RESULT**
</div><!-- .left -->
<div class="right">
$$
PAQ=\left[
  \begin{array}{c|c}
    I_{d\times d} & O_{d\times(n-d)}\\
    \hline O_{(m-d)\times d} & O_{(m-d)\times(n-d)}
  \end{array}
\right]
$$
</div><!-- .left -->
</div>

#Equalities and Inequalities

```{r, echo = FALSE}
equalities = tribble(
  ~desription,                ~cmd,    ~asis,
  "Equality or assignment",   "=",     TRUE,
  "Inequality",               "neq",   FALSE,
  "Less than",                "<",     TRUE,
  "Less than or equal to",    "leq",   FALSE,
  "Greater than",             ">",     TRUE,
  "Greater than or equal to", "geq",   FALSE,
  "Approximately equal to",   "simeq", FALSE,
  "Identical to",             "equiv", FALSE)
  
equalities = equalities %>% 
  by_row(function(x) put_codechunk(x$cmd, x$asis), 
         .collate = "rows", .to = "latex") %>% 
  by_row(function(x) put_result(x$cmd, x$asis), 
         .collate = "rows", .to = "result") %>% 
  select(-asis, -cmd)

knitr::kable(equalities)
```

#Subset and Functions

```{r, echo = FALSE}
set_operations = tribble(
  ~desription,            ~cmd,
  "A is included in B",   "A \\subset B",
  "A includes B",         "A \\supset B",
  "A intersection B",     "A \\cap B",
  "A union B",            "A \\cup B", 
  "x is in A",            "x \\in A",
  "A contains x",         "A \\ni x",
  "x is not in A",        "x \\notin A",
  "for some x",           "\\exists x",
  "for any x",            "\\forall x",
  "tends to / maps into", "\\to",
  "gets",                 "\\gets",  # I've never used this...
  "A implies B",          "A \\implies B")

set_operations = set_operations %>% 
  by_row(function(x) put_codechunk(x$cmd, TRUE), 
         .collate = "rows", .to = "latex") %>% 
  by_row(function(x) put_result(x$cmd, TRUE), 
         .collate = "rows", .to = "result") %>% 
  select(-cmd)

knitr::kable(set_operations)
```

# Mathematical fonts

```{r, echo = FALSE}
math_fonts = tribble(
  ~desription,                                      ~cmd,
  "Set of complex numbers",                        "\\mathbb{C}",
  "Set of rational numbert",                       "\\mathbb{Q}",
  "Set of real numbers",                           "\\mathbb{R}",
  "Expected value of a random variable",           "\\mathbb{E}", 
  "Lagrangian",                                   "\\mathcal{L}",
  "Hamiltonian",                                  "\\mathcal{H}")

math_fonts = math_fonts %>% 
  by_row(function(x) put_codechunk(x$cmd, TRUE), 
         .collate = "rows", .to = "latex") %>% 
  by_row(function(x) put_result(x$cmd, TRUE), 
         .collate = "rows", .to = "result") %>% 
  select(-cmd)

knitr::kable(math_fonts)
```



# Others

**LaTeX Code**
<div class="twocol">
<div class="left">
```
$$
\mathrm{sgn}:S_{n}\to\{-1,1\}
$$ 
```
**RESULT**
</div><!-- .left -->
<div class="right">
$$
\mathrm{sgn}:S_{n}\to\{-1,1\}
$$ 
</div><!-- .left -->
</div>

**LaTeX Code**
<div class="twocol">
<div class="left">
```
$$
  \mathrm{sgn}(\sigma)=\begin{cases}
  -1 & \mbox{if }\sigma\mbox{ is odd}\\
  +1 & \mbox{if }\sigma\mbox{ is even}.
  \end{cases}
$$
```
**RESULT**
</div><!-- .left -->
<div class="right">
$$
  \mathrm{sgn}(\sigma)=\begin{cases}
  -1 & \mbox{if }\sigma\mbox{ is odd}\\
  +1 & \mbox{if }\sigma\mbox{ is even}.
  \end{cases}
$$
</div><!-- .left -->
</div>









